---
title: "Course Project 1"
output:
  html_document:
    fig_height: 4
    highlight: pygments
    theme: spacelab
  pdf_document: default
---

#Load Required Packages

```{r}
library(dplyr)
```

#Download and Load Data

```{r}
url1 = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
if(!file.exists("activity.zip")) {download.file(url1, destfile = "activity.zip")}
unzip("activity.zip")
activity <- read.csv("activity.csv")
```

#Histogram of Total Steps

```{r, warning = F}
#calculate total number of steps
total.steps <- data.frame(with(activity, tapply(steps, date, sum, na.rm = T)))
colnames(total.steps) <- c("Sum")
#plot histogram
hist(total.steps$Sum, breaks = 10, xlab = "Total Number of Steps",
     main = "Histogram of Total Steps", col = "steelblue")
```

#Mean and Median of Steps

```{r}
#mean number of steps
mean.s <- activity %>%
  group_by(as.factor(date)) %>%
  summarise(mean.step = mean(steps, na.rm = T))
print(mean.s)
#median number of steps
median.s <- activity %>%
  group_by(as.factor(date)) %>%
  summarise(median.step = median(steps, na.rm = T))
print(median.s)
```

#Time-Series Plot of Average Steps

```{r}
#average steps across different intervals
mean.s.daily <- data.frame(with(activity, tapply(steps, interval, mean, 
                                                 na.rm = T)))
plot(mean.s.daily, type = "l", xlab = "Time Interval", ylab = "Average Steps",
     main = "Average Steps Across Time")
```

#Interval With the Most Steps

```{r}
colnames(mean.s.daily) <- make.names("mean.steps.daily")
#sort of maximum steps
max.steps <- mean.s.daily[which.max(mean.s.daily$mean.steps.daily), ]
```

The maximum number of steps was at `r max.steps` with interval 835.

#Total Rows With NAs

```{r}
#sum number of NAs
na.rows <- sum(!complete.cases(activity))
```

The total number of NAs in the data was `r na.rows`.

#Replace NAs With Mean Values In New Dataset

```{r}
#create new dataset
n.activity <- activity
#replace NAs by mean values
n.activity$steps <- with(n.activity, 
                       ifelse(is.na(steps), 
                              ave(steps, interval, 
                                  FUN = function(x) mean(x, na.rm = T)), 
                              steps))
```

#Histogram of New Dataset

```{r}
#calculate total number of steps
total.steps.n <- data.frame(with(n.activity, tapply(steps, date, sum, 
                                                  na.rm = T)))
colnames(total.steps.n) <- c("Sum")
#plot histogram
hist(total.steps.n$Sum, breaks = 10, xlab = "Total Number of Steps",
     main = "Histogram of Total Steps With New Dataset", col = "steelblue")
```

The histogram looks more unimodal now with most of the total steps being around 10000.

#Mean and Median Steps of New Dataset

```{r}
#mean number of steps
mean.s.n <- n.activity %>%
  group_by(as.factor(date)) %>%
  summarise(mean.step = mean(steps, na.rm = T))
print(mean.s.n)
#median number of steps
median.s.n <- n.activity %>%
  group_by(as.factor(date)) %>%
  summarise(median.step = median(steps, na.rm = T))
print(median.s.n)
```

The beginning values of mean steps is now a number, instead of being not a number (NaN). The median number of steps begin to have values other than just zero. NAs seems to affect this dataset a lot by skewing the dataset.

#Create Factor Variables of Weekdays and Weekends

```{r}
#convert date to date format
n.activity$date <- as.POSIXct(strptime(n.activity$date, "%Y-%m-%d"))
#create factor variable for weekday and weekend
n.activity$wd <- ifelse(weekdays(n.activity$date, abbreviate = T) ==
                          c("Fri", "Sat", "Sun"), "weekend", "weekday")
```

#Plot Average Steps for Weekday and Weekend

```{r}
n.split <- (split(n.activity, n.activity$wd))
mean.weekday <- tapply((n.split$weekday)$steps, (n.split$weekday)$interval,
                       mean)
mean.weekend <- tapply((n.split$weekend)$steps, (n.split$weekend)$interval,
                       mean)
par(mfrow = c(1, 2))
plot(mean.weekday, type = "l", ylab = "Number of Steps",
     xlab = "Interval", main = "Steps on Weekdays", ylim = c(0, 250))
plot(mean.weekend, type = "l", ylab = "Number of Steps",
     xlab = "Interval", main = "Steps on Weekends", ylim = c(0, 250))
```